---
title: "Maryland Poison Center (MPC) data"
author: "Leah Jager"
date: "6/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(pdftools)
```

### Reading in data from a single file: Allegany County, 2018

Do I want to use `pdf_data()` or `pdf_text()`?  Let's see what each does for my file.  First, using `pdf_text()`:
```{r}
myData <- pdf_text("Allegany County Statistical Report 2018.pdf")
myData
myData[[1]]

myData2 <- pdf_data("Allegany County Statistical Report 2018.pdf")
myData2
myData2[[1]]
```
Both options separate the text from the two pages of the file, which is ideal since I am initially only going to pull the data from the first page.  If I use `pdf_text()` I will have to split on the newline character in order to separate the rows of data, but with `pdf_data()` it is already split to the individual words.  I do want to keep lines of words together, such as "< 12 months", etc for the ages and "Healthcare Facility" for the site, but I think I can do this after `pdf_data()` by using the (x,y) location variables for the text boxes:
```{r}
pdfData <- pdf_data("Allegany County Statistical Report 2018.pdf")
p1Data <- pdfData[[1]]

lineData <- p1Data %>% 
  group_by(y) %>% 
  arrange(x, .by_group=TRUE) %>%
  summarize(line = paste(text, collapse=" "))

lineData
```

Now I still have the problem of the fact that there are really two columns worth of data on this page of the pdf, which is reflected in row 5 of this line data frame, where we see there are 529 total human exposures and 24 animal exposures, both recorded on the same line of the pdf:
```{r}
lineData[5,]
```

I wonder if I can first split the data into two groups by column, and then group by the y value:
```{r}
p1Data %>% group_by(x) %>% arrange(y, .by_group=TRUE) %>% as.data.frame()
```

The end of the first column is at y=258, so I will create a new column to designate left/right column.  Then I can group by this left/right column before grouping by y and summarizing by collapsing across x:
```{r}
p1Data <- p1Data %>%
  mutate(column=ifelse(x < 260, "Left", "Right"))

lineData <- p1Data %>% 
  group_by(column,y) %>% 
  arrange(x, .by_group=TRUE) %>%
  summarize(line = paste(text, collapse=" "))
```

This is looking really good!  However the header of the pdf is causing trouble, since it bridges the two columns.  It will be easier if I remove it, but first I want to extract the county name and the year:

```{r}
orderedData <- p1Data %>% arrange(y,x)
year <- orderedData$text[1]
county <- paste(orderedData$text[4:5],collapse=" ")
county = gsub(",","",county)
```

Now I will remove these header rows and split into two columns.  I want to remove the first 3 rows of the data, which occur up to position y=159.
```{r}
p1Data %>% group_by(y) %>% summarize(n=n())

p1Data <- p1Data %>% filter(y > 160)

lineData <- p1Data %>% 
  group_by(column,y) %>% 
  arrange(x, .by_group=TRUE) %>%
  summarize(line = paste(text, collapse=" "))

lineData %>% as.data.frame()
```

Now I don't think I actually want to collapse across all words in a row, I want to keep the last word as the value and the previous words as the variable name:

```{r}
groupedData <- p1Data %>% 
  group_by(column,y) %>% 
  arrange(x, .by_group=TRUE) %>%
  mutate(type = ifelse(x==max(x), "value", "name"))

countData <- groupedData %>% 
  group_by(column, y) %>%
  arrange(x, .by_group=TRUE) %>%
  summarize(variable = paste(text[type=="name"], collapse=" "), count=text[type=="value"])

countData %>% as.data.frame()

countData <- countData %>% filter(count != "Calls", count!="exposure", count!="Site", count!="Outcome")

countData <- t(countData)

myRow <- as.data.frame(t(as.numeric(countData$count)))
names(myRow) <- countData$variable
myRow$Year <- year
myRow$County <- county

myRow
```

Now to make this into a function that is given the pdf file and returns the row of data:

```{r}

#pdf.file <- "Allegany County Statistical Report 2018.pdf"

pdfMPC.page1 <- function(pdf.file) {
  require(dplyr)
  require(pdftools)
 
# read in the pdf document; select the first page 
pdfData <- pdf_data(pdf.file)
p1Data <- pdfData[[1]]

# get the year and country from the header
orderedData <- p1Data %>% arrange(y,x)
year <- orderedData$text[1]
county <- paste(orderedData$text[4:5],collapse=" ")
county = gsub(",","",county)

# remove the header
p1Data <- p1Data %>% filter(y > 160)

# create the column variable (Left/Right)
p1Data <- p1Data %>%
  mutate(column=ifelse(x < 260, "Left", "Right"))

# group the data by column and height on the page
# keep the last entry of that column/height as the value
# assign the remaining entries for that column/height the name
groupedData <- p1Data %>% 
  group_by(column,y) %>% 
  arrange(x, .by_group=TRUE) %>%
  mutate(type = ifelse(x==max(x), "value", "name"))

# collapse the entries for name together to create the variable name
# keep the value as the count
countData <- groupedData %>% 
  group_by(column, y) %>%
  arrange(x, .by_group=TRUE) %>%
  summarize(variable = paste(text[type=="name"], collapse=" "), count=text[type=="value"])

# remove the rows that aren't variables/counts
countData <- countData %>% filter(count != "Calls", count!="exposure", count!="Site", count!="Outcome")

# create the data frame for this county/date
myRow <- as.data.frame(t(as.numeric(countData$count)))
names(myRow) <- countData$variable
myRow$Year <- year
myRow$County <- county

return(myRow)
}
```

Try out this function on the original dataset and two more files:
```{r}
pdfMPC.page1("Allegany County Statistical Report 2018.pdf")
pdfMPC.page1("Prince Georges County Statistical Report 2018.pdf")
pdfMPC.page1("Talbot County Statistical Report 2018.pdf")
```

I get warnings with both of these additional files.  I expected the issue with Price George's County, because there is an additional blurb at the top that reads "NOTE: This report reflects only the calls to the Maryland Poison Center from Prince Georges County. For complete statistics regarding Prince Georges County, statistics from the National Capitol Poison Center should also be consulted." I am surprise this happed with Talbot, since at a glad the structure of the file looks the same.  But the location data must not actually be the same, and this is the issue with hard coding amounts to determine the header and columns.  I need to look at both of these pdfs through `pdff_data()` to see how different they look.

