---
title: "Maryland Poison Center (MPC) data"
author: "Leah Jager"
date: "6/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

This file contains code to:

* Download pdf documents containing call reports by county and year to the Maryland Poison Center from [https://www.mdpoison.com/](https://www.mdpoison.com/).  Call reports from all 24 Maryland county equivalents for the years of 2006 - 2018 are downloaded.  A sample pdf document can be found [here](https://www.mdpoison.com/media/SOP/mdpoisoncom/factsandreports/reports/countypdf2018/Allegany%20County%20Statistical%20Report%202018.pdf).
* Extract the data from the first page of each of these documents and organize into a data frame.
* Combine the data from each document into a single data frame.
* Clean and recode data to produce a tidy data file; a code book is included at the end.

```{r packages}
library(tidyverse)
library(pdftools)
```

Function to read/organize in the data from a single pdf document into a data frame.  This function take a pdf document name, `pdf.file`, and a file number `i` and returns a data frame with one row containing the county/year data from the corresponding pdf document.
```{r readFunction}
pdfMPC.page1 <- function(pdf.file, i) {
  require(dplyr)
  require(pdftools)
 
# read in the pdf document; select the first page 
pdfData <- pdf_data(pdf.file)
p1Data <- pdfData[[1]]

# get the year and country from the header
year <- p1Data %>% arrange(y,x) %>% slice(1) %>% select(text) %>% as.numeric()

county <- p1Data %>% group_by(y) %>%
  arrange(x, .by_group=TRUE) %>% 
  summarize(line = paste(text, collapse=" ")) %>%
  slice(2) %>% select(line) %>% as.character()

# remove the header 
# the first 3 lines for most, the first 6 lines for PG and M counties
if (county=="Prince Georges County, MD" | county=="Montgomery County, MD") {
  y.cut <- p1Data %>% group_by(y) %>% arrange(y) %>% summarize(n=n()) %>%
    slice(6) %>% select(y) %>% as.numeric()
  p1Data <- p1Data %>% filter(y > y.cut + 1)
} else {
  if (county=="Queen Anne’s" | county=="St. Mary’s") {
      y.cut <- p1Data %>% group_by(y) %>% arrange(y) %>% summarize(n=n()) %>%
    slice(4) %>% select(y) %>% as.numeric()
  p1Data <- p1Data %>% filter(y > y.cut + 1)
  county <- paste0(county," County, MD")
  } else {
    y.cut <- p1Data %>% group_by(y) %>% arrange(y) %>%
      summarize(n=n()) %>%
      slice(3) %>% select(y) %>% as.numeric()
    p1Data <- p1Data %>% filter(y > y.cut + 1)
  }
}

# create the column variable (Left/Right)
p1Data <- p1Data %>%
  mutate(column=ifelse(x < 265, "Left", "Right"))

# group the data by column and height on the page
# keep the last entry of that column/height as the value
# assign the remaining entries for that column/height the name
groupedData <- p1Data %>% 
  group_by(column,y) %>% 
  arrange(x, .by_group=TRUE) %>%
  mutate(type = ifelse(x==max(x) & x==min(x), "name", ifelse(x==max(x), "value", "name")))

# collapse the entries for name together to create the variable name
# keep the value as the count
countData <- groupedData %>% 
  group_by(column, y) %>%
  arrange(x, .by_group=TRUE) %>%
  summarize(variable = paste(text[type=="name"], collapse=" "), count=ifelse(is_empty(text[type=="value"])==FALSE, text[type=="value"],"0"), xmin=min(x))

# remove the any variables/counts that include any part of "Maryland Center"
countData <- countData %>% filter(count != "Maryland", count!="Center",  variable!="Maryland", variable!="Center", variable!="Types of", variable!="Reason for")

## Designate the "Other Reason" variable to say "Other Reason"
countData$variable[countData$variable=="Other" & countData$xmin==min(countData$xmin)] <- "Other Reason"

y.age.min <- countData$y[countData$variable=="Total human exposures"]
y.age.max <- countData$y[countData$variable=="Unintentional"]
y.un.min <- countData$y[countData$variable=="Unintentional"]
y.un.max <- countData$y[countData$variable=="Intentional"]
y.int.min <- countData$y[countData$variable=="Intentional"]
y.int.max <- countData$y[countData$variable=="Other Reason"]
y.other.min <- countData$y[countData$variable=="Other Reason"]

y.ms.min <- countData$y[countData$variable=="Management"]
y.ms.max <- countData$y[countData$variable=="Medical"]
y.mo.min <- countData$y[countData$variable=="Medical"]

countData <- countData %>%
  mutate(variableSub=ifelse(column=="Left" & y > y.age.min & y < y.age.max, "Age:",
         ifelse(column=="Left" & y > y.un.min & y < y.un.max, "Unintentional:",
         ifelse(column=="Left" & y > y.int.min & y < y.int.max, "Intentional:",
         ifelse(column=="Left" & y > y.other.min, "Other:",
         ifelse(column=="Right" & y > y.ms.min & y < y.ms.max, "ManageSite:",
         ifelse(column=="Right" & y > y.mo.min, "MedOutcome:", "")))))))

# remove the rows that aren't variables/counts
countData <- countData %>% filter(count != "Calls", count!="exposure", count!="Site", count!="Outcome", variable!="Medical", variable!="Outcome")

countData <- countData %>%
  mutate(name=paste0(variableSub, variable))

# create the data frame for this county/date
myRow <- as.data.frame(t(as.numeric(gsub(",","",countData$count))))
names(myRow) <- countData$name
myRow$Year <- year
myRow$County <- county
myRow$fileNo <- i #to help diagnose

return(myRow)
}
```

Create a link for each county/year document as well as a file name to store the document in:
```{r createLinks}
countyNames <- c("Allegany County", "Anne Arundel County", "Baltimore City", "Baltimore County", "Calvert County", "Caroline County", "Carroll County", "Cecil County", "Charles County", "Dorchester County", "Frederick County", "Garrett County", "Harford County", "Howard County", "Kent County", "Montgomery County", "Prince Georges County", "Queen Annes County", "Somerset County", "St Marys County", "Talbot County", "Washington County", "Wicomico County", "Worcester County")
years <- 2006:2018

links <- NULL
files <- NULL
for (i in years) {
  for (j in countyNames) {
    countyNameForLink <- paste(unlist(strsplit(j, " ")), collapse="%20")
    if (i != 2016) {
      tempLink <- paste0("https://www.mdpoison.com/media/SOP/mdpoisoncom/factsandreports/reports/countypdf",i,"/",countyNameForLink,"%20Statistical%20Report%20",i,".pdf")} else {
      tempLink <- paste0("https://www.mdpoison.com/media/SOP/mdpoisoncom/factsandreports/reports/county-pdf-",i,"/",countyNameForLink,"%20Statistical%20Report%20",i,".pdf")}
    tempFile <- paste0(j," Statistical Report ", i,".pdf")
    links <- c(links, tempLink)
    files <- c(files, tempFile)
  }
}
```

Download and save documents in a `data` subfolder:
```{r downloadDocuments, eval=FALSE}
for (i in 1:length(links)) {
  download.file(links[i], paste0("./data/",files[i]))
}
```

Extract data from page 1 of each pdf document using the function `pdfMPC.page1()` and bind data from different documents together using `bind_rows()`:
```{r}
d1 <- pdfMPC.page1(paste0("./data/",files[1]),1)
myData <- d1

#options(warn=2)  # to see warnings in the for loop
for (i in 2:length(files)) {
  di <- pdfMPC.page1(paste0("./data/",files[i]),i)
  myData <- bind_rows(myData,di)
  #print(i) # for debugging purposes
}

dim(myData)
names(myData)
```

Recode variables as needed to account for typos in variable names; combine variables that are measuring the same thing but have slightly different names.

Recode age variables to combine the three different versions for `Unknown` together.  Also combine the two versions of `Unknown adult` together.
```{r recodeAge}
myData <- myData %>% 
  mutate(`Age:Unknown`=ifelse(!is.na(`Age:Unknown`), `Age:Unknown`, `Age:Unknown age`)) %>% 
  select(-`Age:Unknown age`) %>%
  mutate(`Age:Unknown`=ifelse(!is.na(`Age:Unknown`), `Age:Unknown`, `Age:Unknown Age`)) %>% 
  select(-`Age:Unknown Age`)

myData <- myData %>% 
  mutate(`Age:Unknown adult`=ifelse(!is.na(`Age:Unknown adult`), `Age:Unknown adult`, `Age:Unknown Adult`)) %>% 
  select(-`Age:Unknown Adult`)
```

Recode management site variables to combine the `Other` and `Unknown` variables into a single `Other/Unknown` variable.  Then combine this new variable with the the two other versions for `Other/Unknown` together.
```{r}
myData <- myData %>% 
  rowwise() %>% 
  mutate(`ManageSite:temp`=ifelse(!is.na(`ManageSite:Other`) | !is.na(`ManageSite:Unknown`), sum(`ManageSite:Other`,`ManageSite:Unknown`, na.rm=TRUE),NA)) %>% 
  select(-`ManageSite:Other`, -`ManageSite:Unknown`) %>% 
  ungroup()

myData <- myData %>% 
  mutate(`ManageSite:Other/Unknown`=ifelse(!is.na(`ManageSite:Other/Unknown`), `ManageSite:Other/Unknown`, `ManageSite:Other/unknown`)) %>%
  select(-`ManageSite:Other/unknown`) %>%
  mutate(`ManageSite:Other/Unknown`=ifelse(!is.na(`ManageSite:Other/Unknown`), `ManageSite:Other/Unknown`, `ManageSite:temp`)) %>% 
  select(-`ManageSite:temp`)
```

Recode other reason variables to combine the `Contamination/Tampering` and `Contamination/tampering` variables.  Also combine the `Other` and `Unknown` variables into a single `Other/Unknown` variable.  Then combine this new variable with the the six other versions for `Other/Unknown` together.
```{r otherRecode}
myData <- myData %>% 
  mutate(`Other:Contamination/Tampering`=ifelse(!is.na(`Other:Contamination/Tampering`), `Other:Contamination/Tampering`, `Other:Contamination/tampering`)) %>%
  select(-`Other:Contamination/tampering`)

myData <- myData %>% 
  rowwise() %>% 
  mutate(`Other:temp`=ifelse(!is.na(`Other:Other`) | !is.na(`Other:Unknown`), sum(`Other:Other`,`Other:Unknown`, na.rm=TRUE),NA)) %>% 
  select(-`Other:Other`, -`Other:Unknown`) %>%
  ungroup()

myData <- myData %>% 
  mutate(`Other:Other/Unknown`=ifelse(!is.na(`Other:Other/Unknown`), `Other:Other/Unknown`, `Other:temp`)) %>% 
  select(-`Other:temp`) %>%
  mutate(`Other:Other/Unknown`=ifelse(!is.na(`Other:Other/Unknown`), `Other:Other/Unknown`, `Other:Other?Unknown`)) %>% 
  select(-`Other:Other?Unknown`) %>% 
  mutate(`Other:Other/Unknown`=ifelse(!is.na(`Other:Other/Unknown`), `Other:Other/Unknown`, `Other:Other/Unknown Reason`)) %>% 
  select(-`Other:Other/Unknown Reason`) %>% 
  mutate(`Other:Other/Unknown`=ifelse(!is.na(`Other:Other/Unknown`), `Other:Other/Unknown`, `Other:Unknown Reason`)) %>% 
  select(-`Other:Unknown Reason`) %>% 
  mutate(`Other:Other/Unknown`=ifelse(!is.na(`Other:Other/Unknown`), `Other:Other/Unknown`, `Other:Unknown/Other`)) %>% 
  select(-`Other:Unknown/Other`)
```

Recode intentional reason variables to combine the `Bite or sting` and `Bite or Sting` variables.  Also recode to combine the `Food poisoning` and `Food Poisoning` variables. Also combine the three versions of `Other/Unknown` and `Unknown` variables together.
```{r intentionalRecode}
myData <- myData %>% 
  mutate(`Unintentional:Bite or sting`=ifelse(!is.na(`Unintentional:Bite or sting`), `Unintentional:Bite or sting`, `Unintentional:Bite or Sting`)) %>%
  select(-`Unintentional:Bite or Sting`)

myData <- myData %>% 
  mutate(`Unintentional:Food poisoning`=ifelse(!is.na(`Unintentional:Food poisoning`), `Unintentional:Food poisoning`, `Unintentional:Food Poisoning`)) %>%
  select(-`Unintentional:Food Poisoning`)

myData <- myData %>% 
  mutate(`Unintentional:Other/Unknown`=ifelse(!is.na(`Unintentional:Other/Unknown`), `Unintentional:Other/Unknown`, `Unintentional:Unkknown`)) %>%
  select(-`Unintentional:Unkknown`) %>%
  mutate(`Unintentional:Other/Unknown`=ifelse(!is.na(`Unintentional:Other/Unknown`), `Unintentional:Other/Unknown`, `Unintentional:Unknown`)) %>%
  select(-`Unintentional:Unknown`)
```

Order the data in a meaningful way to bring the `fileNo`, `County`, and `Year` variables to the front of the data frame and put the age groups in a more meaningful order.
```{r orderVariables}
myData <- myData %>%
  select(sort(names(myData)))

myData <- myData %>%
  select(fileNo, County, Year, `Total human exposures`, `Animal Exposures`, `Information Calls`, `Age:< 12 months`, `Age:1 year`, `Age:2 years`, `Age:3 years`, `Age:4 years`, `Age:5 years`, `Age:6-12 years`, `Age:13-19 years`, `Age:20-59 years`, `Age:>60 years`, everything())
```

Clean the data further by:

* Fix the typo for Caroline County for 2010+, where the `>60 years` category was incorrectly coded as `<60 years`.
* Combine the `20-59 years` and `>60 years` categories into an `Adult` category for 2010 on.  (For 2006-2009, there was no age breakdown for adults, just an `Adult` category.  This new `Adult` category for the later years will be comparable.  But keep the age breakdowns for these later years as well.)
* Fix the typo that the 2015 data for Sommerset County is mis-labeled 2014 in the pdf document.
* Incorporate the `Unknown Adult` group into the `Adult` category for the 2010+ years, but also keep this category separate for when using the age breakdown

```{r dataCleaning}
myData <- myData %>% 
  mutate(`Age:>60 years`=ifelse(!is.na(`Age:>60 years`), `Age:>60 years`, `Age:<60 years`)) %>% 
  select(-`Age:<60 years`)

myData <- myData %>% 
  mutate(`Age:Adult`=ifelse(Year < 2010, `Age:Adult`, `Age:>60 years`+`Age:20-59 years`))

myData$Year[myData$fileNo==235] <- 2015

myData <- myData %>% 
  mutate(`Age:Adult`=ifelse(Year < 2010, `Age:Adult`, ifelse(!is.na(`Age:Unknown adult`),`Age:>60 years`+`Age:20-59 years` + `Age:Unknown adult`, `Age:>60 years`+`Age:20-59 years`)))
```

Finally, make the variable names easier to work with by removing symbols and spaces and switching to camel case:
```{r renameVariables}
namesNew <- names(myData)
namesNew <- gsub(pattern=":",replacement="_", namesNew)
namesNew <- gsub(pattern=" or s",replacement="/S", namesNew)
namesNew <- gsub(pattern="/",replacement=".", namesNew)
namesNew <- gsub(pattern="<",replacement="Under", namesNew)
namesNew <- gsub(pattern=">",replacement="Over", namesNew)
namesNew <- gsub(pattern="years",replacement="yr", namesNew)
namesNew <- gsub(pattern="year",replacement="yr", namesNew)
namesNew <- gsub(pattern="months",replacement="mo", namesNew)
namesNew <- gsub(pattern=" p",replacement=" P", namesNew)
namesNew <- gsub(pattern=" s",replacement=" S", namesNew)
namesNew <- gsub(pattern=" a",replacement=" A", namesNew)
namesNew <- gsub(pattern=" c",replacement=" C", namesNew)
namesNew <- gsub(pattern=" h",replacement=" H", namesNew)
namesNew <- gsub(pattern=" e",replacement=" E", namesNew)
namesNew <- gsub(pattern="Site.non",replacement="Site.Non", namesNew)
namesNew <- gsub(pattern=" ",replacement="", namesNew)

names(myData) <- namesNew
```

Check the data names and dimensions:
```{r dataCheck}
names(myData)
dim(myData)
```

Write this clean data to a .csv file:
```{r writeData}
write_csv(myData, "MPCdataFINAL.csv")
```

